Trends of SDR in Thessaloniki

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 11:10

For completeness, we will provide below a brief overview of the clear
sky identification algorithm (CSid), along with the site specific
thresholds. To calculate the reference clear sky
\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
optimization process, as described by \citet{Long2000} and
\citet{Reno2016}. The target of the optimization was the minimization of
a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
algorithmic function ``optimise'', which is an implementation based on
the work of \citet{Brent1973}, from the library ``stats'' of the R
programming language \citep{RCT2023}. The optimization and the selection
of the clear sky reference model, was performed on SDR observations for
the period 2016 - 2021. During the optimization, eight simple clear sky
radiation models were tested (Daneshyar-Paltridge-Proctor,
Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by \citet{Reno2012} and
evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
with the factor \(a = 0.965\) yields one of the lowest root mean squared
errors (RMSE), while the procedure, manages to characterize the majority
of the data. Thus, our clear sky reference is derived by the
Eq.~\ref{eq:ahau}. \begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation} where: \(n\) is the total number of daylight records,
\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
by CSid, \(a\) is a hypothetical adjustment factor, and
\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
radiation models. \begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar
zenith angle (SZA).

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below. A data point is flagged
as ``clear-sky'' if all criteria are satisfied, otherwise it is
considered to be ``cloud-sky''. Each criterion was applied for a running
window of \(11\) consecutive one-minute measurements, and the
characterization is assigned to the central value of the window. Each
parameter, was calculated both from the observations and the reference
clear sky model, for each comparison. The allowable range of variation
is defined by the model-derived value of the parameter multiplied by a
factor plus an offset. The factors and the offsets were determined
empirically, by manual inspecting each filters performance on selected
days, and adjusting them accordingly, during an iterative process.

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
  \ref{eq:MeanVIP}). \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75
  < M_{\text{}i} <
  1 \times M_{\text{CSref},i} + 75
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points
  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation} \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation} where: \(t_i\) is the time each SDR measurement has
  been measured
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
  \(11\) sequential points, normalized by the mean
  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather} For this criterion, \(\sigma_i\) should be below a
  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance
  and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather} For this criterion, \(X_i\) should be below a certain
  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
    X_i < 7.5 \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

Due to a significant measurement uncertainty near the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all one-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
the Sun's intensity variability, using a time series of satellite TSI
observations. The TSI data we use are part of the ``NOAA Climate Data
Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. Where
the initial daily values, were interpolated to match with the time step
of our measurements. The final dataset contains \(6589967\) one-minute
measurements, of which, \(84.2\%\) were identified as under clear-sky
conditions and subsequently \(15.8\%\) as under cloud-sky conditions.

In order to investigate the SDR trends, we implemented an appropriate
aggregation scheme to the 1-minute data to derive a series in coarser
timescale. To preserve the representativeness of the data we used the
following criteria: a) for the daily mean values we accept days with
more than 50\% of the daytime measurements, present and valid, b)
monthly values were computed from daily means only when at least 20 days
were available. To create the daily and monthly climatological means, we
averaged the data based on the day of year and calendar month,
respectively. For the seasonal means we averaged the mean daily values
in each season (Winter: December - February, Spring: March - May, etc.).
Finally, each data set was deseasonalized by subtracting the
corresponding climatological annual cycle (daily or monthly) from the
actual data. To estimate SZA contribution to the SDR trends, the
one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
the morning and afternoon hours, and then were deseasonalized as
mentioned above.

>>>>

For completeness, we will provide below a brief overview of the clear
sky identification algorithm (CSid), along with the site specific
thresholds. To calculate the reference clear sky
\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
optimization process, as described by \citet{Long2000} and
\citet{Reno2016}. The target of the optimization was the minimization of
a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
algorithmic function ``optimise'', which is an implementation based on
the work of \citet{Brent1973}, from the library ``stats'' of the R
programming language \citep{RCT2023}. The optimization and the selection
of the clear sky reference model, was performed on SDR observations for
the period 2016 - 2021. During the optimization, eight simple clear sky
radiation models were tested (Daneshyar-Paltridge-Proctor,
Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by \citet{Reno2012} and
evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
with the factor \(a = 0.965\) yields one of the lowest root mean squared
errors (RMSE), while the procedure, manages to characterize the majority
of the data. Thus, our clear sky reference is derived by the
Eq.~\ref{eq:ahau}. \begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation} where: \(n\) is the total number of daylight records,
\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
by CSid, \(a\) is a hypothetical adjustment factor, and
\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
radiation models. \begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar
zenith angle (SZA).

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below. A data point is flagged
as ``clear-sky'' if all criteria are satisfied, otherwise it is
considered to be ``cloud-sky''. Each criterion was applied for a running
window of \(11\) consecutive one-minute measurements, and the
characterization is assigned to the central value of the window. Each
parameter, was calculated both from the observations and the reference
clear sky model, for each comparison. The allowable range of variation
is defined by the model-derived value of the parameter multiplied by a
factor plus an offset. The factors and the offsets were determined
empirically, by manually inspecting each filter's performance on
selected days and adjusting them accordingly during an iterative
process. The criteria are:

\begin{enumerate}
\def\labelenumi{\alph{enumi})}
\tightlist
\item
  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
  \ref{eq:MeanVIP}). \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75\,Wm^{-2}
  < M_{i} <
  1 \times M_{\text{CSref},i} + 75\,Wm^{-2}
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points
  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation} \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation} where: \(t_i\) is the time stamp of each SDR
  measurement.
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
  \(11\) sequential points, normalized by the mean
  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather} For this criterion, \(\sigma_i\) should be below a
  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance
  and the change in clear sky irradiance over each measurement interval.
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather} For this criterion, \(X_i\) should be below a certain
  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
    X_i < 7.5 \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

Due to the significant measurement uncertainty when the Sun is near the
horizon, we have excluded all measurements with SZA greater than
\(85^\circ\). Moreover, due to some obstructions around the site (hills
and buildings), we excluded data with azimuth angle between \(35^\circ\)
and \(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically not visible from the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all one-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
the Sun's intensity variability, using a time series of satellite TSI
observations. The TSI data we used are part of the ``NOAA Climate Data
Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. The
initial daily values of this dataset were interpolated to match the time
step of our measurements. In the final dataset \(6589967\) one-minute
measurements, of which, \(84.2\%\) were identified as under clear-sky
conditions and \(15.8\%\) as under cloud-sky conditions.

In order to investigate the SDR trends, we implemented an appropriate
aggregation scheme to the 1-minute data to derive a series in coarser
time-scales. To preserve the representativeness of the data we used the
following criteria: a) for the daily mean values we accept days with
more than 50\% of the daytime measurements present and valid, b) monthly
values were computed from daily means only when at least 20 days were
available. To create the daily and monthly climatological means, we
averaged the data based on the day of year and calendar month,
respectively. For the seasonal means we averaged the mean daily values
in each season (Winter: December - February, Spring: March - May, etc.).
Finally, each data set was deseasonalized by subtracting the
corresponding climatological annual cycle (daily or monthly) from the
actual data. To estimate the SZA effect on the SDR trends, the
one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
the morning and afternoon hours, and then were deseasonalized as
mentioned above.
